---
title: "Workshop SVEPM 2022"
author: "Eleftherios Meletis & Aurélien Madouasse"
date: '2022-03-23'
site: distill::distill_website
description: |
  Substantiating freedom from infection -  available statistical methods to prove freedom from infection in an output-based framework
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(htmltools.preserve.raw = FALSE)
```

```{r meta, echo=FALSE}
library(metathis)
meta() %>%
  meta_general(
    description = "Substantiating freedom from infection -  available statistical methods to prove freedom from infection in an output-based framework",
    generator = "xaringan and remark.js"
  ) %>% 
  meta_name("github-repo" = "LefMel/SVEPM_2022_wk")
```


# Workshop Agenda

- 09:00 - 09:30 Introduction (framing the problem, Terminology: Se/Sp/risk factors, prevalence)
- 09:30 - 10:00 Scenario tree analysis (scenario tree notation & calculations, simple example)
 - 10:00 - 10:30 Coffee break
- 10:30 - 11:00 Bayesian hierarchical true prevalence estimation (apparent/true prevalence estimation with JAGS/Bayesian inference, explain full hierarchical model)
- 11:00 - 11:45 STOCfree model
- 11:45 - 12:00 Wrap-up & Closure

# Slides

- Introduction [Slides](https://LefMel.github.io/SVEPM_2022_wk/01_concepts/disease_freedom_concepts)
- Scenario tree method [Slides](https://htmlpreview.github.io/?https://github.com/LefMel/SVEPM_2022_wk/blob/main/02_Scenario_tree_method/Scenario-tree-method.html)
- Bayesian hierarchical true prevalence estimation [Slides](https://htmlpreview.github.io/?https://github.com/LefMel/SVEPM_2022_wk/blob/main/03_Bayesian_Hierarchical/bayes_prev.html)
- STOCfree model [Slides](https://LefMel.github.io/SVEPM_2022_wk/04_STOCfree/STOCfree_model)
- Conclusions [Slides](https://htmlpreview.github.io/?https://github.com/LefMel/SVEPM_2022_wk/blob/main/05_conclusion/Conclusions.html)
