---
title: "Substantiating freedom from disease as a statistical problem"
author: "Aurélien Madouasse & Eleftherios Meletis"
date: "2022-03-23"
output:
  xaringan::moon_reader:
    css: ["default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: inverse, center, middle

# Objective of this presentation

---
# Objective

- Frame the problem of substantiating freedom from disease as a statistical problem

---
class: inverse, center, middle

# Sampling for proving freedom from disease

---
## Sampling for proving freedom from disease

- Evidence of infection sought by testing animals from the population of interest

--

 - Testing all the animals is usually too costly $\rightarrow$ sampling
 
--

 - $\geq 1$ (true) positive $\rightarrow$ infection is present
 
--

 - 0 test positive $\rightarrow$ infection is absent OR infection is present, but no infected included in the sample

--


- Most of the time, the problem consists in **quantifying the probability that the infection is not present given that it has not been detected**

--

- The question can be formulated as: 

 - **how sure are we that the infection is truly absent given that we did not find it?** or

 - **what is the probability of a false negative surveillance outcome?**

---
## Exercise 1

- Consider a large population in which 20% of animals are infected

- 10 animals of this population are tested with a perfect test

--

- What are the probabilities of getting 0 to 10 positives among 10 animals tested?

 - What is the distribution to use?
 - Which parameters?

--

- Assuming a large population, the probability of getting $Y = k$ positive animals out of $n$ animals tested when the disease prevalence is $\pi$ follows a Binomial distribution:

$$p(Y = k) \sim Binomial(n, \pi) $$

---
## Exercise 1

- Consider a large population in which 20% of animals are infected

- 10 animals of this population are tested with a perfect test

- What are the probabilities of getting 0 to 10 positives among 10 animals tested?

 - Write the R code to display and plot the results

--

```{r}
dbinom(0:10, 10, .2)
```

---
## Exercise 1

- Consider a large population in which 20% of animals are infected

- 10 animals of this population are tested with a perfect test

- What are the probabilities of getting 0 to 10 positives among 10 animals tested?

 - Write the R code to display and plot the results

```{r, fig.height = 4}
barplot(dbinom(0:10, 10, .2))
```

---
## Exercise 1


```{r, fig.height = 4}
barplot(dbinom(0:10, 10, .2),
        names.arg = 0:10,
        xlab = "Number of test positives out of 10 animals tested")
```

---
## Exercise 2

- Compare what happens when:

 - 10 animals sampled / disease prevalence is 20%
 - 20 animals sampled / disease prevalence is 20%
 - 10 animals sampled / disease prevalence is 1%
 - 20 animals sampled / disease prevalence is 1%

---
## Exercise 2

```{r, fig.keep='none'}
par(mfrow = c(2, 2))
# n = 10, pi = 20%
barplot(dbinom(0:10, 10, .2),
        names.arg = 0:10,
        xlab = "Number of test positives out of 10 animals tested",
        main = "Prevalence = 20%")
# n = 20, pi = 20%
barplot(dbinom(0:10, 20, .2),
        names.arg = 0:10,
        xlab = "Number of test positives out of 20 animals tested",
        main = "Prevalence = 20%")
# n = 10, pi = 1%
barplot(dbinom(0:10, 10, .01),
        names.arg = 0:10,
        xlab = "Number of test positives out of 10 animals tested",
        main = "Prevalence = 1%")
# n = 20, pi = 1%
barplot(dbinom(0:10, 20, .01),
        names.arg = 0:10,
        xlab = "Number of test positives out of 20 animals tested",
        main = "Prevalence = 1%")
```

---
## Exercise 2

```{r, echo = FALSE, fig.width = 10}
par(mfrow = c(2, 2))
barplot(dbinom(0:10, 10, .2),
        names.arg = 0:10,
        xlab = "Number of test positives out of 10 animals tested",
        main = "Prevalence = 20%")

barplot(dbinom(0:10, 20, .2),
        names.arg = 0:10,
        xlab = "Number of test positives out of 20 animals tested",
        main = "Prevalence = 20%")

barplot(dbinom(0:10, 10, .01),
        names.arg = 0:10,
        xlab = "Number of test positives out of 10 animals tested",
        main = "Prevalence = 1%")

barplot(dbinom(0:10, 20, .01),
        names.arg = 0:10,
        xlab = "Number of test positives out of 20 animals tested",
        main = "Prevalence = 1%")
```

---
## Exercise 3

- What are the probabilities of (falsely) concluding that disease is absent when:
 - 10 animals sampled / disease prevalence is 20%
 - 20 animals sampled / disease prevalence is 20%
 - 10 animals sampled / disease prevalence is 1%
 - 20 animals sampled / disease prevalence is 1%

--

```{r}
dbinom(0, 10, .2)
dbinom(0, 20, .2)
dbinom(0, 10, .01)
dbinom(0, 20, .01)
```

---
## Key messages

- Probability of a false negative surveillance outcome $\nearrow$ when:

 - sample size $\searrow$
 
 - prevalence $\searrow$
 
 